---
title: "Assignment 2 - Categorical Data"
author: "Michael Francis"
output: pdf_document
classoption: landscape
---


```{r setup, include=FALSE, warnings=FALSE, message=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)


library(readxl)
library(tidyverse)
library(jsonlite)
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(grid)

# read in shots data from a practice on 10-11-2019, using in-out device and apple watch
data <- read_csv("C:/Users/mgffr/OneDrive/Desktop/MSDS 455/MSDS-455-Data-Visualization/Assignment 2/shots.csv")

#net and midline at 0,0
#serve box is (-21, 0) to (-21, 13.5), (-21, 0) to (-21, -13.5), (21, 0) to (21, 13.5), (21, 0) to (21, -13.5),
#baseline at (-39, 13.5) to (-39, -13.5) and (39, 13.5) to (39, -13.5)

out_bounds <- data.frame(
  x = c(-39, 39, 39, -39, -39),
  y = c(18, 18, -18, -18, 18)
)

t_lines <- data.frame(
  x = c(21, 21, 21, -21, -21, -21),
  y = c(-13.5, 13.5, 0, 0, -13.5, 13.5)
)

court <- ggplot() +
  geom_path(data = out_bounds, aes(x = x, y = y)) +
  geom_path(data = t_lines, aes(x = x, y = y)) +
  geom_path(aes(x = c(-39, 39), y = c(13.5, 13.5))) + # lower singles lines
  geom_path(aes(x = c(-39, 39), y = c(-13.5, -13.5))) + # upper singles lines
  geom_path(aes(x = c(0,0), y = c(18, -18)), lty = 2) + # net line
  ylim(c(-20, 20)) + xlim(c(-50, 50)) +
  #scale_color_fivethirtyeight() +
  theme_fivethirtyeight() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

michael <- subset(data, data$playingPlayerID=="Michael")
michael$in_out <- ifelse(michael$in_out==1, "In", "Out")

p1 <- ggplot(michael, aes(x=stroke, y=ballSpeed, fill=in_out)) +
  geom_boxplot() +
  scale_color_fivethirtyeight() +
  theme_fivethirtyeight() +
  ggtitle("Ball Speed by Stroke and In/Out") +
  theme(axis.title = element_text(), 
        plot.title = element_text(hjust = 0.5)) + 
  ylab('Avg Ball Speed') +
  xlab('Game')  + labs(color='') 

p2 <- court +
  geom_point(data = michael,
             aes(x = x, y = y, colour = stroke)
  ) +
  theme(legend.position = "bottom", 
        plot.title = element_text(hjust = 0.5)) +
  ggtitle("Tennis Stroke Bounce Positions")  + 
  labs(color='Ball Speed')

```

# Tennis Serve Speed During a Match

```{r plot, echo=FALSE, fig.height=6.5, fig.width=15, warnings=FALSE, message=FALSE}
grid.arrange(p1, p2, ncol=2)
```

\begin{flushright}
Source: Francis, M (2020, July 17). [Tennis Singles Match] Retrieved from https://github.com/MichaelFrancis23/MSDS-455-Data-Visualization/blob/master/Assignment\%202/shots.csv
\end{flushright}

- Data was taken from an In/Out tennis camera on 10/11/2019 for one of my tennis practices using a ball machine.
- 1st serves have a higher speed than 2nd serves, forehands have a higher speed than backhands, speed is more varied for shots hit out.
- More serves than ground strokes are hit into the net, majority of shots hit to the right side of the court.
