---
title: "Assignment 1 - Temporal Data"
author: "Michael Francis"
output: pdf_document
classoption: landscape
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(readxl)
library(tidyverse)
library(jsonlite)
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(grid)

### UPDATE THIS ####
date <- "2019-10-12"


inOut <- fromJSON(txt = "C:/Users/mgffr/OneDrive/Desktop/Tennis/InOut/Singles/20191012-123330.json",flatten=T)

#net and midline at 0,0
#serve box is (-21, 0) to (-21, 13.5), (-21, 0) to (-21, -13.5), (21, 0) to (21, 13.5), (21, 0) to (21, -13.5),
#baseline at (-39, 13.5) to (-39, -13.5) and (39, 13.5) to (39, -13.5)


out_bounds <- data.frame(
  x = c(-39, 39, 39, -39, -39),
  y = c(18, 18, -18, -18, 18)
)

t_lines <- data.frame(
  x = c(21, 21, 21, -21, -21, -21),
  y = c(-13.5, 13.5, 0, 0, -13.5, 13.5)
)

court <- ggplot() +
  geom_path(data = out_bounds, aes(x = x, y = y)) +
  geom_path(data = t_lines, aes(x = x, y = y)) +
  geom_path(aes(x = c(-39, 39), y = c(13.5, 13.5))) + # lower singles lines
  geom_path(aes(x = c(-39, 39), y = c(-13.5, -13.5))) + # upper singles lines
  geom_path(aes(x = c(0,0), y = c(18, -18)), lty = 2) + # net line
  ylim(c(-20, 20)) + xlim(c(-50, 50)) +
  scale_color_fivethirtyeight() +
  theme_fivethirtyeight() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())


shots <- inOut$shots
shots$x <- as.numeric(shots$x)
shots$y <- as.numeric(shots$y)
shots$playingx <- as.numeric(shots$playingx)
shots$playingy <- as.numeric(shots$playingy)
shots$receivingx <- as.numeric(shots$receivingx)
shots$receivingy <- as.numeric(shots$receivingy)


shots$Time <- NA
shots$Time <- as.POSIXct(as.numeric(shots$timestamp), origin="1970-01-01", tz="UTC") 


shots$TimeSinceLast <- NA
for(i in 1:nrow(shots)){
  if(i==1){
    shots$TimeSinceLast[i] <- 0
  }else{
    shots$TimeSinceLast[i] <- as.numeric(difftime(shots$Time[i], shots$Time[i-1], units = "secs"))
  }
}

head(shots[order(-shots$TimeSinceLast),c(23,24)], 15)
serves <-subset(shots, shots$lineCallType!=0) 
servesNew <- serves[,c(6,13,23,24)]
x <- head(serves[order(-serves$TimeSinceLast),c(23,24)], 15)


#warm-up & 1, 2&3, 4&5, 6&1, 2&3, 4&5, 6&7
#109 (98)   , 178, 213, 231, 290,    , 
#                      (255)

# warmup 97, game 1 108 (michael = inOut$playerID)
# game 2 147, game 3 177 (michael != inOut$playerID)
# game 4 204, game 5 230 (michael = inOut$playerID)
# game 6 254, game 1 289 (michael != inOut$playerID)
# game 2 333, game 3 368 (michael = inOut$playerID)
# game 4 388, game 5 417 (michael != inOut$playerID)
# game 6 443, game 7 rest (michael = inOut$playerID)

rm(serves, servesNew); gc()

shotsTest <- shots

for(i in 1:nrow(shots)){
  if(shots$playingPlayerID[i]==inOut$playerID & i<=108){
    shots$playingPlayerID[i] <- "Michael"
  }else if(shots$playingPlayerID[i]!=inOut$playerID & i<=177 & i>108){
    shots$playingPlayerID[i] <- "Michael"
  }else if(shots$playingPlayerID[i]==inOut$playerID & i<=230 & i>177){
    shots$playingPlayerID[i] <- "Michael"
  }else if(shots$playingPlayerID[i]!=inOut$playerID & i<=289 & i>230){
    shots$playingPlayerID[i] <- "Michael"
  }else if(shots$playingPlayerID[i]==inOut$playerID & i<=368 & i>389){
    shots$playingPlayerID[i] <- "Michael"
  }else if(shots$playingPlayerID[i]!=inOut$playerID & i<=417 & i>368){
    shots$playingPlayerID[i] <- "Michael"
  }else if(shots$playingPlayerID[i]==inOut$playerID & i<=nrow(shots) & i>417){
    shots$playingPlayerID[i] <- "Michael"
  }else{
    shots$playingPlayerID[i] <- "Ben Ginster"
  }
}

for(i in 1:nrow(shots)){
  if(shots$receivingPlayerID[i]==inOut$playerID & i<=108){
    shots$receivingPlayerID[i] <- "Michael"
  }else if(shots$receivingPlayerID[i]!=inOut$playerID & i<=177 & i>108){
    shots$receivingPlayerID[i] <- "Michael"
  }else if(shots$receivingPlayerID[i]==inOut$playerID & i<=230 & i>177){
    shots$receivingPlayerID[i] <- "Michael"
  }else if(shots$receivingPlayerID[i]!=inOut$playerID & i<=289 & i>230){
    shots$receivingPlayerID[i] <- "Michael"
  }else if(shots$receivingPlayerID[i]==inOut$playerID & i<=368 & i>289){
    shots$receivingPlayerID[i] <- "Michael"
  }else if(shots$receivingPlayerID[i]!=inOut$playerID & i<=417 & i>368){
    shots$receivingPlayerID[i] <- "Michael"
  }else if(shots$receivingPlayerID[i]==inOut$playerID & i<=nrow(shots) & i>417){
    shots$receivingPlayerID[i] <- "Michael"
  }else{
    shots$receivingPlayerID[i] <- "Ben Ginster"
  }
}

shots$playingPlayerID <- ifelse(shots$playingPlayerID=="Michael", "Ben Ginster", "Michael")
shots$receivingPlayerID <- ifelse(shots$receivingPlayerID=="Michael", "Ben Ginster", "Michael")

shots$in_out <- ifelse(shots$in_out==2, 1, shots$in_out)

shots$Game <- ifelse(as.numeric(row.names(shots)) <=97, 0, ifelse(
  as.numeric(row.names(shots)) <=108, 1, ifelse(
    as.numeric(row.names(shots)) <=147, 2, ifelse(
      as.numeric(row.names(shots)) <=177, 3, ifelse(
        as.numeric(row.names(shots)) <=204, 4, ifelse(
          as.numeric(row.names(shots)) <=230, 5, ifelse(
            as.numeric(row.names(shots)) <=254, 6, ifelse(
              as.numeric(row.names(shots)) <=289, 7, ifelse(
                as.numeric(row.names(shots)) <=333, 8, ifelse(
                  as.numeric(row.names(shots)) <=368, 9, ifelse(
                    as.numeric(row.names(shots)) <=388, 10, ifelse(
                      as.numeric(row.names(shots)) <=417, 11, ifelse(
                        as.numeric(row.names(shots)) <=443, 12, 13)))))))))))))






michael <- subset(shots, shots$playingPlayerID=="Michael")

michaelServes <- subset(michael, michael$lineCallType!=0) 
michaelShots <- subset(michael, michael$lineCallType==0) 


michaelServesByGame <- michaelServes %>%
  select(Game, in_out, ballSpeed) %>%
  group_by(Game, in_out) %>%
  summarise(AvgBallSpeed = mean(as.numeric(ballSpeed), na.rm=T))





michaelServesByGame$in_out <- ifelse(michaelServesByGame$in_out==1, "In", "Out")

michaelServes$ballSpeed <- as.numeric(michaelServes$ballSpeed)






p1 <- ggplot(michaelServesByGame, aes(x=factor(Game), y=AvgBallSpeed, group=in_out, color=in_out)) +
  geom_line(aes(color=in_out)) +
  geom_point(aes(color=in_out)) +
  scale_color_fivethirtyeight() +
  theme_fivethirtyeight() +
  ggtitle("Serve Speed by Service Game") +
  theme(axis.title = element_text(), 
        plot.title = element_text(hjust = 0.5)) + 
  ylab('Avg Ball Speed') +
  xlab('Game')  + labs(color='') 

p2 <- court +
  geom_point(data = michaelServes,
             aes(x = x, y = y, colour = ballSpeed)
  ) +
  theme(legend.position = "bottom", 
        plot.title = element_text(hjust = 0.5)) +
  ggtitle("Bounce Positions")  + 
  labs(color='Ball Speed')  + 
  scale_color_gradient(low="blue", high="red") 

```

# Tennis Serve Speed During a Match

```{r pressure, echo=FALSE, fig.height=6.5, fig.width=15}
grid.arrange(p1, p2, ncol=2)
```

\begin{flushright}
Source: Francis, M (2020, July 9). [Tennis Singles Match] Retrieved from https://github.com/MichaelFrancis23/MSDS-455-Data-Visualization/blob/master/Assignment\%201/20191012-123330.json
\end{flushright}

- Data was taken from an In/Out tennis camera on 10/12/2019 for one of my tennis matches, net balls are not captured.
- 2nd serves tend to be slower than 1st serves but are hit in more often. This corresponds with the average lower speed of serves called in.
- As a match progresses players tire and serve speed decreases
